KPIs and Visualisations for report

### 2022-08-01 Report/KPIs and Visualisations for report.Rmd

---
title: "NGO analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed and avoiding
# catastrophic dashboard errors if they're not available
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using the fontawesome package for more up-to-date (and currently supported)
# Fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

Sidebar {.sidebar}
=======================================================================

### File upload

Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.

Then upload an Excel file in the set format. <!--# Need to insert link when repository is public -->

Click 'Analyse' to analyse the data.

```{r sidebar}
# HTML upload
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel upload
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Analyse button
actionButton("click", "Analyse")
```

If desired, a copy of the analysis in PDF format may then be downloaded.

```{r sidebar_button}
# download button - using uiOutput() because of an apparent bug
# button and report are created in the second reactive chunk
uiOutput("downloadUI")
```

Dashboard
=======================================================================

Row 
-----------------------------------------------------------------------

### KPI Fundraising
```{r}
valueBoxOutput("KPI_box_fundraising")
```

### KPI OPX
```{r}
valueBoxOutput("KPI_box_OPX")
```

### KPI Private growth
```{r}
valueBoxOutput("KPI_box_private")
```

### KPI Public growth
```{r}
valueBoxOutput("KPI_box_public")
```

### KPI Private/public proportion
```{r}
valueBoxOutput("KPI_box_proportion")
```

Row
-----------------------------------------------------------------------

### Net income
```{r}
plotlyOutput("arsresultat_plotly")
```

### Total equity/liabilities
```{r}
plotlyOutput("balance_plotly")
```

Row 
-----------------------------------------------------------------------

### Revenues
```{r}
plotlyOutput("revenue_plotly")
```

### Expenses
```{r}
plotlyOutput("expenses_plotly")
```

Row 
-----------------------------------------------------------------------

### Transparency: CEO compensation
```{r}
plotlyOutput("CEO_plotly")
```

Details
=======================================================================

Row {data-height=60}
-----------------------------------------------------------------------

### NGO analysed

```{r}
# Using htmlOutput() to use html styling
htmlOutput("title")
```

Row 
-----------------------------------------------------------------------

@@

The KPIs are based on the following calculations, using data derived from the Excel sheet:

@@

@@

@@

@@

Row 
-----------------------------------------------------------------------

### Graphs

The dashboard includes five graphs, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

```{r reactive_title_value_boxes, include=FALSE}
# Shiny actions begin here =====================================================
# First reactive chunk =========================================================

# wait for button click
observe({

  # Excel data frame ===========================================================

  ## Validating and creating long version of Excel file, if possible------------
  excel_test <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))

    widen_excel_test(input$excel$datapath)
  })

  ## Creating the primary Excel data frame--------------------------------------
  exceldata <- reactive({
    widen_excel(excel_test())
  })

  # Value Boxes ================================================================

  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark"
  # if the value is NA

  ## Fundraising ROI------------------------------------------------------------
  # Creating Fundraising ROI value
  KPI_fundraising_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    }
  })

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box_fundraising <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_fundraising_num()),
      caption = "Fundraising ROI",
      # Colour depends on value
      color = if (is.na(KPI_fundraising_num())) {
        "primary"
      } else if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_fundraising_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-donate"
      }
    )
  })

  ## OPX KPI--------------------------------------------------------------------
  # Creating OPX KPI
  KPI_OPX_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        # Calculating the value as a %
        transmute(
          KPI_OPX =
            sum(revenue_private_contributions,
              revenue_investments, assets_cash_in_bank_and_cash_equivalents,
              na.rm = TRUE
            ) /
              sum(liability_short_term_grants_payable,
                liability_revocable_endowments,
                liability_long_term_grants_payable,
                na.rm = TRUE
              ) * 100
        ) |>
        # Just the single number
        pull() |>
        first()
    }
  })

  # Creating value box for OPX KPI
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      # colour depends on value
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      # icon depends on value
      icon = if (is.na(KPI_OPX_num())) {
        "fa-solid fa-circle-xmark"
      } else if (KPI_OPX_num() < 100) {
        "fa-solid fa-scale-unbalanced"
      } else {
        "fa-solid fa-scale-balanced"
      }
    )
  })

  ## Private donations growth---------------------------------------------------
  # Creating private donations growth
  KPI_private_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_private_contributions[1] -
              revenue_private_contributions[2]) /
              revenue_private_contributions[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      # colour depends on value
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 0) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_private_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-hands-helping"
      }
    )
  })

  ## Public grants growth-------------------------------------------------------
  # Creating public grants growth
  KPI_public_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_public_grants[1] - revenue_public_grants[2]) /
              revenue_public_grants[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      # colour depends on value
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 0) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_public_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-building-columns"
      }
    )
  })

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # calculating the value as a %
        transmute(
          proportion =
            sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) /
              sum(revenue_public_grants, revenue_private_contributions,
                revenue_membership_fees,
                na.rm = TRUE
              ) * 100
        ) |>
        pull() |>
        # getting the most recent year
        first()
    }
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      # Adding in the same check as elsewhere - colours match if all are NA
      color = if (is.na(KPI_proportion())) {
        "primary"
      } else {
        "info"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_proportion())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-chart-pie"
      }
    )
  })

  ## Excel data test -----------------------------------------------------------
  # (also below, in the second chunk)
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  # the button click
}) |> bindEvent(input$click)
```

```{r reactive_report_plots, include=FALSE}
# Beginning of the second reactive chunk =======================================
# It's necessary to split the observers into separate parts, with all necessary
# code duplicated, because of R's limitations.

observe({

  # Data frames ================================================================
  ## Creating data frame from html file-----------------------------------------
  html_currency <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))

    # Testing and creating data frame
    html_test_create(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  htmldata <- reactive({
    read_proff_two(html_currency())
  })

  # Recreating, to make available to this chunk as well
  ## Validating and creating long version of Excel file, if possible------------
  excel_test <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))

    widen_excel_test(input$excel$datapath)
  })

  ## Creating the primary Excel data frame--------------------------------------
  exceldata <- reactive({
    widen_excel(excel_test())
  })

  # NGO title ==================================================================
  NGO <- reactive({
    # checking for the required column again
    if ("firm" %in% names(excel_test())) {
      # fetching the name
      excel_test() |>
        # Turning the needed column into a list
        pull(firm) |>
        # discarding all NAs in case the name is somewhere
        purrr::discard(is.na) |>
        # choosing the first available name
        first()
    } else {
      validate("Please upload data using the specified Excel template.")
    }
  })

  title <- reactive({
    # testing if NGO() is NA
    if (is.na(NGO())) {
      # Printing warning if NA
      "Name of NGO missing"
      # Printing name of NGO
    } else {
      NGO()
    }
  })

  # Creating NGO title using html formatting (colour based on MetBrewer
  # "Hokusai2")
  output$title <- renderText({
    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
  })

  # Report Output ==============================================================

  ## Download button -----------------------------------------------------------
  output$downloadUI <- renderUI({
    if (available_excel() && available_html()) {
      downloadButton("report", "Download report")
    }
  })

  ## Data exists condition testing----------------------------------------------
  # Essentially pretty random tests to connect to the individual
  # validate() functions upstream in the reactive chain

  # Excel
  available_excel <- reactive({
    TRUE
  }) |>
    bindEvent(setequal(
      names(exceldata()),
      c("year", "firm", "category", "sub_category", "name", "value")
    ))

  # HTML
  available_html <- reactive({
    TRUE
  }) |> bindEvent("lederlonn" %in% names(htmldata()))

  ## Report download and parameters --------------------------------------------
  output$report <- downloadHandler(

    # creating the file name using the NGO title and system date
    filename = function() {
      str_c(title(), "_", Sys.Date(), ".pdf")
    },
    content = function(file) {
      # Based on code in Mastering Shiny -- creating a temporary directory and
      # file
      tempReport <- file.path(tempdir(), "report.Rmd")
      file.copy("report.Rmd", tempReport, overwrite = TRUE)

      # parameters
      params <- list(
        htmlfile = input$html$datapath,
        excelfile = input$excel$datapath,
        start = input$start,
        end = input$end
      )

      # notification -- disappears when report generation is complete
      id <- showNotification(
        "Generating report...",
        duration = NULL,
        closeButton = FALSE
      )
      on.exit(removeNotification(id), add = TRUE)

      # R Markdown render
      rmarkdown::render(tempReport,
        output_file = file,
        params = params,
        envir = new.env(parent = globalenv())
      )
    }
  )

  # Plots ======================================================================

  ## Annual result--------------------------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # renaming for plotly hover
        rename(
          Year = year,
          `Net income` = arsresultat,
          `Net positive` = annual_profit
        ) |>
        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Net income (in thousands)"
        )
    }
  })

  # Creating plotly output
  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  ## Balance--------------------------------------------------------------------
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # rename for plotly hover
        rename(
          Year = year,
          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
          `Positive equity` = balance_positive
        ) |>
        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Total equity/liabilities (in thousands)",
        )
    }
  })

  # Creating plotly output
  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  ## Expenses-------------------------------------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

  # Creating plotly output
  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot())
  )

  ## Revenue--------------------------------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`,
              levels =
                c(
                  "Other", "Commercial activities",
                  "Operational activities", "Investments",
                  "Membership fees", "Private contributions",
                  "Public grants"
                )
            )
        ) |>
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

  # Creating plotly output
  output$revenue_plotly <- renderPlotly(
    ggplotly(revenue_plot())
  )

  ## CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    # Testing for the presence of data
    if (sum(!is.na(htmldata()$lederlonn)) == 0 &&
      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      # Creating plot if any data exist
      # Note that the plot will still look empty, but have a legend and scales,
      # if all values are explicit zeroes
      htmldata() |>
        # preparing for pivoting
        select(
          year, lederlonn, leder_annen_godtgjorelse
        ) |>
        # Compensation in ordinary numbers
        mutate(
          lederlonn = lederlonn * 1000,
          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
        ) |>
        # renaming for the purposes of the plot
        rename(
          Year = year,
          Salary = lederlonn,
          `Other compensation` = leder_annen_godtgjorelse
        ) |>
        # pivoting to join the two values to use fill
        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
        ggplot(aes(Year, value, fill = Compensation)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # pretty colours
        scale_fill_met_d("Hokusai2", direction = -1) +
        labs(
          x = "", y = "Total compensation", title = "CEO compensation"
        )
    }
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  # Reactive functions =========================================================

  ## Currency scale-------------------------------------------------------------
  # Reactive function to check for currency and use currency code or ?
  currency_scale <- reactive({
    scale_y_continuous(
      # Checking for uniform currency code
      labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
        # Using the currency code if uniform
        label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
        # alternative prefix if multiple currencies in the data
        label_dollar(prefix = "? ")
      )
    )
  })

  ## Plotly caption-------------------------------------------------------------
  # Reactive function to create plotly caption (ggplot captions and subtitles
  # are lost in the conversion) (shamelessly stolen from the Internet, with
  # minor modifications).
  # Checks for uniform currency code and prints a warning
  # if there are multiple currency codes.
  plotly_caption <- reactive({
    list(
      x = 1, y = -0.1,
      text = ifelse(length(unique(html_currency()$valutakode)) == 1,
        "", "Note: not all values are denominated in the same currency"
      ),
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 11)
    )
  })

  ## scale maximum--------------------------------------------------------------
  # Reactive function to create the maximum value for revenue and expenses plots
  # the tests are probably not necessary, but included to be on the safe side
  maximum <- reactive({
    if (datatest() == TRUE || setequal(
      names(exceldata()),
      c("year", "firm", "category", "sub_category", "name", "value")
    )) {
      NA
    } else {
      excel_test() |>
        # removing the firm and sub-category columns
        select(-firm, -sub_category) |>
        # filtering to retain only revenue and expenses entries
        filter(category == "Revenue" | category == "Expenses") |>
        # grouping and summarising to calculate maximum value
        group_by(year, category) |>
        summarise(
          sum = sum(value, na.rm = TRUE)
        ) |>
        # Creating list of sums
        pull() |>
        # selecting the highest value
        max()
    }
  })

  ## Excel data test (also above)-----------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)
```

```{r non-reactive functions, include=FALSE}
# Data reading functions =======================================================

## Functions to read html file -------------------------------------------------
html_test_create <- function(file, start, end) {
  # Reading file and node
  htmltest <- read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")

  ### A series of tests to catch problems before they spit out errors-----------
  # Checking if node exists
  if (class(htmltest) == "xml_node") {
    htmltable <- htmltest |>
      html_table()
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Basic check if the table is correct
  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
    htmlready <- htmltable
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Testing if the correct years have been entered
  if (as.character(all_of(start)) %in% names(htmlready) &&
    as.character(all_of(end)) %in% names(htmlready)) {
    # Creating the first part of the data frame
    read_proff_one(htmltable, start, end)
  } else {
    validate("Please ensure you have entered the correct start and end years.")
  }
}

### First html transformation function, invoked by html_test_create() ----------
# (stopping to allow checks for differing currency codes)
read_proff_one <- function(object, start, end) {
  object |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column to avoid having to deal with NAs later
    fill(
      valutakode,
      .direction = "downup"
    )
}

### Second html transformation function, creating the data frame ---------------
read_proff_two <- function(object) {
  object |>
    # removing superfluous currency column
    select(-valutakode) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots (red/green profit/loss)
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

## Functions to read Excel file ------------------------------------------------
### First Excel transformation -------------------------------------------------
# creating df used by NGO title function and maximum scale function
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

### Second Excel function - creating primary Excel data frame-------------------
widen_excel <- function(object) {
  # testing if all columns are present
  if (setequal(
    names(object),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    object |>
      # uniting category and name to deal with multiple 'Other'
      unite(
        entry, c(category, name)
      ) |>
      # deleting Sub-category and Firm
      select(-sub_category, -firm) |>
      # pivoting to create 'tidy' variables
      pivot_wider(names_from = entry, values_from = value) |>
      # cleaning names
      clean_names()
    # Notification if all columns are not present
  } else {
    validate("Please upload data using the specified Excel template.")
  }
}

# Support functions ============================================================

## KPI_to_string ---------------------------------------------------------------
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
  # avoid passing an NA to subsequent operations.)
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}

## scale_fill_green ------------------------------------------------------------
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}

## no_information plot ---------------------------------------------------------
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  empty_plot("No information available")
}

## bad_data plot ---------------------------------------------------------------
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  empty_plot("Something is wrong\nwith the Excel data.")
}

## zero_data_plot --------------------------------------------------------------

zero_data <- function() {
  empty_plot("All values are zero.\nThis is unusual and may\nindicate a data problem.")
}

## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    # annotating the plot with the specified warning message
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

>>>>

---
title: "KPIs and Visualisations for report"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-08-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

On the basis of such data as we could gather we developed five potential KPIs:

@@

This KPI is not a true measure of fundraising ROI, as it only includes (as we understand OPX's data) the direct fundraising expenses, and does not include major indirect expenses such as salary costs.

***

@@

This KPI was suggested by OPX.

***

@@

***

@@

This KPI may say something about the NGO's sources of funding. There are prima facie no specific desirable funding mix, but this measure may be of interest to some donors.

***

### Visualisations

The dashboard includes five plots, three based on data from Proff.no and two based on data from the Excel sheet.

-   **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
-   **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
-   **Revenues**: Based on the category *Revenue* in the Excel sheet.
-   **Expenses**: Based on the category *Expenses* in the Excel sheet.
-   **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
    -   This appears to increasingly be treated as a necessary measure of transparency and performance, but the values should normally be evaluated in the context of the size of the NGO and the sector(s) within which it operates.

The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.
